Bridging the Gap Between Visual and Analytical Machine Learning Testing

### 2023-09-24 report/esecfse24.tex
feat(report): 2023-09-24 12:26 notes on related work
feat(report): 2023-09-24 18:23 updates

@@

\citeauthor{pimentel2019large} mine 1.4 million notebooks from Github to conduct an empirical study on the good and bad coding practices when developing software using computational notebooks. The authors also make their dataset of notebooks publicly available on Zenodo~\footnote{https://zenodo.org/record/2592524}. Computational notebooks were originally designed to promote sharing reproducible research since they allow presenting information by interweaving natural text, code and other multimedia such as tables, images and videos. However, the authors discover that Jupyter notebooks are far from being reproducible due to bad coding practices such as lack of separation of concern, testing and versioning.

@@

@@

% TODO kandel2012enterprise: semi-structured interview study (n=35) with data analysts from 25 organizations across diverse sectors. Relationship between data analysis and how oganizational features of enterprise affect it. Recurring painpoints, open challenges and guidelines for building better visual analysis research.

% TODO liu2019understanding: 
\subsection{Visualisations and Software Engineering}\label{sec:visualisations}

% bavishi2021vizsmith
% ??

@@

@@

% TODO try using captionof for the listing as well, so that the caption for the listing and figure align at the bottom
\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python, caption=Assertion on the accuracy of the ML model.]
assert accuracy_score(y_test, pred) > 0.95
  \end{lstlisting}
  \label{lst:svm}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-04.png}
  \captionof{figure}{Visualisation of the decision boundary of a SVM.}
  \label{fig:svm}
\end{minipage}

>>>>

Computational notebooks have been ubiquitously adopted by the machine learning community for developing ML enabled systems. Although originally intended to promote reproducible software, computational notebooks are infamous for the exact opposite since they do not follow software engineering best practices.

The process of understanding and cleaning the underlying training data---commonly referred to as "data wrangling"---tends to dominate the ML development workflow~\cite{CITME}. Data wrangling is high iterative. Practitioners often explore several hypotheses, visual designs, methods, tools, algorithms and data sources to arrive at the final implementation. Studies have been conducted to understand how practitioners generate, evaluate and manage such alternatives. ~\cite{liu2019understanding,kandel2012enterprise}.

Developing software in computational notebooks introduces a new paradigm for writing code. In addition to code, notebooks also contain additional information in the form of natural language text and images. To better understand the process of working within computational notebooks and the challenges faced along the way, studies with human subjects have been conducted\cite{head2019managing}.

 ML practitioners often experiment with several versions of code to implement the same functionality. Prior work has explored how practitioners manage the different versions of code and the challenges that they face along the way.

@@

@@

% code quality; reproducibility; best practices...
\citeauthor{pimentel2019large} mine 1.4 million notebooks from Github to conduct an empirical study on the good and bad coding practices when developing software using computational notebooks. The authors also make their dataset of notebooks publicly available on Zenodo~\footnote{https://zenodo.org/record/2592524}. Computational notebooks were originally designed to promote sharing reproducible research since they allow presenting information by interweaving natural text, code and other multimedia such as tables, images and videos. However, the authors discover that Jupyter notebooks are far from being reproducible due to bad coding practices such as lack of separation of concern, testing and versioning.

@@

% DONE kandel2012enterprise: semi-structured interview study (n=35) with data analysts from 25 organizations across diverse sectors. Relationship between data analysis and how oganizational features of enterprise affect it. Recurring painpoints, open challenges and guidelines for building better visual analysis research.

% DONE liu2019understanding: semi-structed interview study (n=12) with data workers with various expertise and domain knowledge. Paper wants to understand the process behind exploring "alternatives": different hypothesis, visual design, methods, tools, algorithms, data sources. The studies explores the motivation behind exploring such alternatives, how alternatives help with sense making process, high-level process of alternatives and how practitioners generate, explore and manage such alternatives.

computational notebooks have been widely adopted by the data science and machine learning community for developing ML models. Unlike traditional software, ML enabled systems evolve through change in not only the code base, but also the complexity of the model and the underlying training data.

Naturally, research is active in understanding the process of developing using notebooks. Prior work has looked into best practices
\subsection{Visualisations and Machine Learning}\label{sec:visualisations}

% TODO bavishi2021vizsmith: paper presents VizSmith, an automatic
% data visualisation framework that accepts data in the form of pandas
% dataframe along with a list of columns the user wishes to visualise.
% The tool then automatically produces visualisations for the
% specified columns.

% TODO amershi2015modeltracker: model building is a cognitively
% disruptive step; practitioners are required to iterate between model
% building and error analysis. Paper presents ModelTracker, a
% visualisation tool that presents all the information in traditional
% summary statistics along with the performance of the model. Tool
% validated with users who used the tool for 6 months in a production
% setting.

% TODO gortler2022neo: Research with ML practitioners at Apple and
% identify traditional confusion matrix does not support complex data
% structures that are prevalent today. Paper proposes Neo, an
% interactive and multi-output confusion matrix.

% TODO yang2022pattern: Paper conducts study on interaction techniques for Multivariate matrix visualisations.

% TODO wang2023gam: paper proposes GAM COACH, a tool that generates customizable counterfactual explanations for Generalized Additive Models (GAMs). The tools uses interactive visualisations to allow users to iteratively generate algorithmic recourse plans. The tool is qualitatively validated by users(n=41).

% TODO ahn2023escape: paper proposes ESCAPE, a visual analytics tool that allows users to identify systematic errors in AI enabled systems. The tool promotes a human-in-the-loop approach to identify spurious associations and concept associated misclassifications.

% TODO wang2023drava: paper proposes DRAVA, a visual analytics tool to relate human concepts with semantic dimensions extracted by ML models during Disentangled representation learning>. The tool also provides feedback to minimize the mismatches and helps practitioners obtain data insights fron concept-driven exploration. The tool is validated using application scenarios in an experimental setting.

% TODO robertson2023angler: authors propose ANGLER, a visual analytics tools to help practitioners prioritise machine translation model improvements. Tool is validated with machine translation experts (n=7). The tool allows users to validate interesting hypotheses for prioritising model improvements.

% TODO bernard2019visual: paper proposes visual-analytics tool to perform data pre-processing of multi-variate time series data.

% TODO gschwandtner2018know: paper proposes visual analytics technique to perform data profiling of time series data

% TODO yuan2021survey: survey of visual analytics techniques used in ML. Paper proposes taxonomy of visual analytics techniques: before model building, during model building and after model building.

% TODO wexler2020if: paper proposes WHAT-IF tool, a model agnostic, interactive visualisation tool the understanding ML models. The tool allows users to visually analyse and verify their hypothesis regarding the training data and how the model works. The tool requires a trained model and a sample dataset. The tool is integrated with Tensorflow, Jupyter and Collaboratory notebooks. The tool allows conterfactual reasoning, investigate decision boundariesa and investigate how change in data affects the prediction.

% TODO hohman2019visual: survey of visual analytics techniques specifically for deep learning models. Visual analytics has been rapidly used for model explanation, interpretation, debugging and improvement.

@@

\lstinputlisting[]{nbformat.json}
\captionof{lstlisting}{Internal JSON structure of Jupyter Notebooks.}
\label{lst:nbformat}

The data comprised within Jupyter Notebooks is stored in the JSON format, and thus can be parsed programmatically. Listing~\ref{lst:nbformat} shows the internal json structure of Jupyter Notebooks, adopted from the official nbformat website.~\footnote{https://nbformat.readthedocs.io/en/latest/index.html}. For simplicity, we only show the parts that are relevant for this study.

Notebooks contain four top level keys namely \texttt{metadata}, \texttt{nbformat}, \texttt{nbformat\_minor} and \texttt{cells}. The \texttt{matadata} key consists of additional information regarding the notebook such as the kernel and programming language. \texttt{nbformat} and \texttt{nbformat\_minor} contain the version of nbformat used in the notebook. The \texttt{cells} key consists of a list of all cells within the notebook.

Listing~\ref{lst:nbformat} presents three examples of cells. Each cell is represented as a dictionary with the \texttt{cell\_type}, \texttt{metadata} and \texttt{source} keys. Cells can be of three types: "markdown", "code" or "raw". The contents of each cell can be found within the \texttt{source} key, as a list of strings.

The output generated by a code cell is stored within the \texttt{outputs} key. Static visualisations are stored as base64 encoded png images under the \texttt{"image/png"} key, which is nested within the \texttt{data} key inside \texttt{outputs}. Since a cell can generate multiple outputs, the \texttt{outputs} key is a list.

@@

% TODO try using captionof for the listing as well, so that the caption for the listing and figure align at the bottom
\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
assert accuracy_score(y_test, pred) > 0.95
  \end{lstlisting}
  \captionof{lstlisting}{Assertion on the accuracy of the ML model.}
  \label{lst:svm}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-04.png}
  \captionof{figure}{Visualisation of the decision boundary of a SVM.}
  \label{fig:svm}
\end{minipage}
